\documentclass{article}

\title{Reverse engineering Lotus 1-2-3 file formats}
\author{Geoffrey Mon (geofbot@gmail.com)}
\date{\today}

\begin{document}
\maketitle

\section{File extensions}
\begin{itemize}
\item \textbf{.wks}: up to at least 1A
\item \textbf{.wk1}: up to at least 2.2
\item \textbf{.wk2}: apparently, some versions of 2.3+?
  interestingly, not supported by SmartSuite 9.8
\item \textbf{.wk3}: apparently, 3.x?
\item \textbf{.wk4}: apparently, 4.x?
\item \textbf{.123}: SmartSuite 9.8, others?
\end{itemize}

\section{Saving files}
When Lotus 1-2-3 9.8 running on WINE saves to a file, if
the new version of the file is shorter
than the old version file, then the old
data at the end of the file is not actually removed
(at least when I remove the contents of one cell).
In fact, an End-of-File record is written to the file at the
appropriate position, leaving the existing data at the end
as corrupted and unread records.
I'm not sure if this is a consequence of using WINE,
or if this is the behavior for all versions of the software.

\textbf{However}, when explicitly overwriting one file with
another file, the old file is completely removed and
the new file is saved in its place.

\section{Quirks}
All of the .wk1
test files from openpreserve/format-corpus
except KSBASE.WK1 have a peculiarity:
their calculation mode records use 0x01 instead of 0x00 or 0xFF;
according to the original repository some of these
files were generated with Quattro Pro;
perhaps this is a quirk of that particular software?

\section{Formats}
\subsection{.wks, .wk1, .wk2}

What's the difference between .wks, .wk1, and .wk2?

.wk1 uses 64-bit double-precision floating point, but
Lotus 1-2-3 2.x and 3.x (according to respective manuals, page 1-17)
only supports values from 1E-99 to 9.99E+99.

\subsection{.wk3}
Supports 256 columns, 8192 rows.

Addresses take the form of row (2 bytes), tab/sheet (1 byte), col (1
byte).

Numbers are encoded as 80-bit long doubles,
although Lotus SmartSuite 9.8 does not support numbers larger
than doubles (~1.8E308).

Header:
                tag: 0x0,
                len: 0x1a,
                dat: [
                    0x0,  file version
                    0x10, file version
                    0x4, file sub (whatever that is)
                    0x0, file sub
                    0x0, active range start row
                    0x0, active range start row
                    0x0, active range start sheet
                    0x0, active range start col
                    0x6, active range end row
                    0x0, active range end row
                    0x0, active range end sheet
                    0x2, active range end col
                    0xe, total revisions
                    0x0, total revisions
                    0x30, major id?
                    0x8d, minor id?
                    0x1,
                    0x2,
                    0x0,
                    0x0,
                    0x0,
                    0x1,
                    0x0,
                    0x0,
                    0x0,
                    0x1,
                ],


\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: xetex
%%% End:
